Online {\IterComp}

### 2017-11-16 motivation.tex
motivation

    \paragraph{Representative input sets are hard to create.} For developers to create artificial input sets, they first need to know how
    their application is actually being used. This is not always easy or even possible. Then, they need to create data that will trigger a
    similar behavior and this needs a deep knowledge of the application and considerable effort. This will have to be a continuous work in
    progress, to handle changes in the way the application is used.

    Collecting real input data online and processing them again offline to evaluate optimizations is not a solution either. It requires
    manual modifications to the application to save the data, more modifications to avoid side effects while reusing data, and even more
    modifications to ensure that the application does the same work every time the same data are reused, despite changes in the rest of
    the system state. This is significant engineering work. Even if this was not the case, many applications are just not able to save
    data for privacy or storage overhead reasons.

@@

    \paragraph{Challenges of using real inputs online.} We could overcome the problems of representative inputs by just evaluating
    optimization sequences online while the application is being used. 
    %Optimization sequences are compared based on how much faster the program becomes.
    This gives us live data but creates a new problems.
    With live data, we cannot just compare runtimes. 
    When all of the optimization sequences are tested against the same inputs, it is just a matter of comparing runtimes.
    A shorter runtime for processing the same data means a more efficient binary.
    With live data, we cannot do this, each binary processes
    whatever the input happens to be. We need another efficiency metric that only requires a single execution of an optimized binary with a
    single input.

    \paragraph{Existing efficiency metrics are not up to the task.} Application-specific efficiency metrics~\cite{alameldeen06,coppa14} may
    work well, but the developer has to choose the appropriate metric and then manually modify the application to calculate it. Speedup
    over a common baseline binary is a good application-agnostic metric, equating efficiency with how much faster a given input is
    processed. It would be ideal, if we did not have to use the same input twice, one with the baseline and one with the optimized binary.

    Typical single-run performance metrics, like runtime or Instructions Per Cycle~(IPC), are not fit for purpose.
    To show this, we performed iterative compilation executing each optimization sequence with a different input and then measuring both runtime and IPC.
    The experiments were performed in the set of inputs \#2 discussed in Figure~\ref{fig:motivation_wrong_inputs}.
    For the iterative compilation guided by runtime, it resulted on an average slowdown of 0.97 compared to \texttt{-O3}.
    For the iterative compilation guided by IPC, it resulted in an average slowdown of 0.51 with an IPC of 3.25., while the optimization expected with an average speedup of 1.07 has an average IPC of 2.31.
    This result agrees with the known fact that higher IPC does not necessarily translate to lower execution time when comparing different versions of the program.

>>>>

    \paragraph{Representative input sets are hard to create.} For developers to create artificial input sets, they first need to know how
    their application is actually being used. This is not always easy or even possible. Then, they need to create data that will trigger a
    similar behavior as when the application is used by users. This needs a deep knowledge of the potential user behavior and
    considerable effort. This will have to be a continuous work in
    progress, to handle changes in the way the application is used.

    Collecting real input data online and processing them again offline to evaluate optimizations is not a solution either. It requires
    manual modifications to the application to save the data, more modifications to avoid side effects while reusing data, and even more
    modifications to ensure that the application does the same work every time the same data are reused, despite changes in the rest of
    the system state. \FIXME{ZW: This is a bit vague. I think we need an example to illustrate why the application may not perform the same
    work if the same input is used.}
    This is significant engineering work. Even if this was not the case, many applications are just not able to save
    data for privacy or storage overhead reasons.

@@

    \paragraph{Challenges of using real inputs online.} We could overcome the problems of representative inputs by just evaluating
    optimization sequences online while the application is being used.
    %Optimization sequences are compared based on how much faster the program becomes.
    This gives us live data but creates a new problems.
    With live data, we cannot just compare runtimes.
    When all of the optimization sequences are tested against the same inputs, it is just a matter of comparing runtimes.
    A shorter runtime for processing the same data means a more efficient binary.
    With live data, we cannot do this, each binary processes
    whatever the input happens to be.
    As an example, considering a sorting algorithm. To measure the efficiency of an optimization for the algorithm, we need to look at not just
    how long does the algorithm run but also how many items it processes per run. Simply comparing the running time of two invocations does not
    tell us which program version is faster, as the two invocations may not process the same number of items. As a result, we need another efficiency metric
    that only requires a single execution of an optimized binary with a single input.

    \paragraph{Existing efficiency metrics are not up to the task.} Application-specific efficiency metrics such as the number of transactions per second may
    work well, but the developer has to choose the appropriate metric and then manually modify the application to calculate it~\cite{alameldeen06,coppa14}. Speedup
    over a common baseline binary is a good application-agnostic metric, equating efficiency with how much faster a given input is
    processed. It would be ideal, if we did not have to use the same input twice, one with the baseline and one with the optimized binary.

    Typical single-run performance metrics, like runtime or Instructions Per Cycle~(IPC), are not fit for purpose.
    To show this, we performed iterative compilation executing each optimization sequence with a different input and then measuring both runtime and IPC.
    The experiments were performed in the set of inputs \#2 discussed in Figure~\ref{fig:motivation_wrong_inputs}.
    For the iterative compilation guided by runtime, it resulted on an average slowdown of 0.97 compared to \texttt{-O3}.
    For the iterative compilation guided by IPC, it resulted in an average slowdown of 0.51 with an IPC of 3.25, while the optimization expected with an average speedup of 1.07 has an average IPC of 2.31.
    This result agrees with the known fact that higher IPC does not necessarily translate to lower execution time when comparing \emph{different versions} of the program.
